### 2010-04-27 PhysicalParameters/abstract.tex
added two abstracts

\documentclass{bmd2010a}

\begin{document}

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  An Accurate Method of Measuring a Bicycle's Physical Parameters\\}
{\fontsize{16pt}{20pt}\selectfont%
  Motorcycle Dynamics Symposium\\}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{Jason K. Moore}\\}
  \textit{Mechanical and Aerospace Engineering\\
          University of California, Davis\\
          One Shields Avenue, Davis, CA 95616, USA\\
          e-mail: jkmoor@ucdavis.edu
}\vspace{10pt}\\
  {\fontsize{12}{14}{Mont Hubbard}\\}
  \textit{Affiliation}\vspace{10pt}\\
  {\fontsize{12}{14}{D. E. Author}\\}
  \textit{Affiliation}\vspace{10pt}\\
\end{flushleft}

\section*{Abstract}
Accurate measurments of a bicycle's physical parameters are required for
realistic dynamic simulations. For the most basic models the geometry, mass,
mass location and mass distributions must be measured. More complex models
require measurements of tire characteristics, stiffness, damping, etc. This
paper concerns itself with the measurement of bicycle parameters required for
simulations of the benchmark bicycle presented in \cite{Meijaard2007}. This
bicycle model is made of four rigid bodies, has ideal rolling, frictionless joints,
and is assumed to be symmetric about the frame center plane. A set of 25
physical parameters are used to describe the geometry, mass, mass location and
mass distribution of each of the rigid bodies. These parameters are presented
relative to a reference frame when the bicycle is in the nominal upright
configuration. The methods described herein are based primarily off of the work
done in \cite{Kooijman2006} but have been refined for improved accuracy and
methodology. \cite{Roland1971} measured an experimental bicycle frame in a
similar fashion and both \cite{Dohring1953} and \cite{Singh1971} have used
similar techniques with scooters. We measure the characteristics of six
different bicycles, two of which are setup in two different configurations, for
a total of eight different parameter sets that can be used with, but not
limited, to the benchmark bicycle model. The parameters set is also complete
for non-linear benchmark bicycle model.

The bicycles were chosen for both variety and convience. The six bicycles are as follows:
\begin{description}
    \item[Batavus Browser] A Dutch style city bicycle. The bicycle was measured
        in two configurations. One with the instrumentation described in
        \cite{Kooijman2009a} and one as from the manufacturer.
    \item[Batavus Stratos Deluxe] A Dutch style city bicycle
    \item[Batavus Crescendo Deluxe] A Dutch style city bicycle
    \item[Gary Fisher Mountain Bike] A hardtail mountain bicycle
    \item[Bianchi Pista] A modern steel frame track racing bicycle
    \item[Yellow Bicycle] A stripped down aluminum frame road bicycle. This
        bike was measured in two configurations, the second being the fork was
        rotated in the headtube 180 degrees for larger trail.
\end{description}
\subsection*{Geometry}
The benchmark bicycle needs five basic geometric measurements: the wheelbase
$w$, steer axis tilt $\lambda$, trail $c$, and the wheel radii $r_\mathrm{F}$
and $r_\mathrm{R}$. The wheelbase and steer axis tilt were measured directly
with a scale and a digital level, respectively. The trail and wheel radii were
measured indirectly by fork offset and loaded wheel circumference.
\subsection*{Mass}
The mass of each of the rigid bodies were measured with a precision scale.
\subsection*{Mass location}
The center's of mass of each of the rigid bodies were measured by hanging the
rigid bodies through their centers of mass and assuming symmetry. The
orientation angle of the headtube angle and the point of suspension were
measured and the center's of mass of the frame and fork were calculated from
these values. The wheel centers of mass were assumed to be in their ideal
geometric center.
\subsection*{Mass Distribution}
The moment's of inertia of the rigid bodies were measured with both a torsional
pendulum and a compound pendulum. Once again the assumed symmetry of the
bicycle was utlized to reduce the amount of measurements needed. The
in-symmetric-plane moments of inertia were calculated from measurements of the
rigid bodies period of oscillation when hung from a torsional pendulum at
different orienatation angles. The out of plane moments of inertia were
measured by swinging the rigid bodies as a compound pendulum.
\subsection*{Results}
Parameter sets for all six bicycles (and all eight
configurations)~\ref{tab:par}  will be
presented, along with basic comparisons of the open loop dynamics of the
different types of bicycles. By examining the mass, damping and stiffness
matrices of the system and the eigenvalues at different speeds anedoctal
descriptions of handling are compared to the dynamics.
\begin{table*}[tb]
\centering
\caption{COMBINED BICYCLE AND RIDER PARAMETER VALUES FOR THE BATAVUS BROWSER.}
{\small
\begin{tabular}{lcl}
&&\\
\hline
Parameter & Symbol & Value \\
\hline
wheel base & $w$ & 1.121 m \\
trail & $c$ & 0.069 m \\
steer axis tilt & \emph{$\lambda$} & 0.400 rad \\
gravity & \emph{g} & 9.81 N kg$^{-1}$ \\
forward speed & \emph{v} & various m s$^{-1}$ \\

\multicolumn{3}{l}{\textbf{R}\emph{ear wheel R}}\\
radius & \emph{$r_\mathrm{R}$} & 0.341 m\\
mass & \emph{$m_\mathrm{R}$} &  3.11 kg \\
mass moments of inertia & \emph{$(I_{\mathrm{R}xx},\
I_{\mathrm{R}yy})$} &
(0.088,\ 0.152) $\mathrm{kg\ m}^2$ \\

\multicolumn{3}{l}{\emph{rear} \textbf{B}\emph{ody and frame B}}\\
position center of mass & \emph{$(x_\mathrm{B},\ z_\mathrm{B})$} &
($0.276$,\ $-0.538$) m\\
mass & \emph{$m_\mathrm{B}$} & 9.86 kg \\
mass moments of inertia & $\left[ \begin{array}{ccc}
I_{\mathrm{B}xx} &  0 & I_{\mathrm{B}xz}\\
0 & I_{\mathrm{B}yy} & 0 \\
I_{\mathrm{B}xz} & 0 & I_{\mathrm{B}zz}
\end{array} \right] $ &
$\left[ \begin{array}{ccc}
0.527 &  0 & -0.114\\
0 & 1.317 & 0 \\
-0.114 & 0 & 0.759
\end{array} \right] \ \mathrm{kg\ m}^{2}$\\

\multicolumn{3}{l}{\emph{front} \textbf{H}\emph{andlebar and fork assembly H}}\\
position center of mass & \emph{$(x_\mathrm{H},\ z_\mathrm{H})$} &
($0.867$,\ $-0.748$) m\\
mass & \emph{$m_\mathrm{H}$} & 3.22 kg \\
mass moments of inertia & $\left[ \begin{array}{ccc}
I_{\mathrm{H}xx} &  0 & I_{\mathrm{H}xz}\\
0 & I_{\mathrm{H}yy} & 0 \\
I_{\mathrm{H}xz} & 0 & I_{\mathrm{H}zz}
\end{array} \right] $ &
$\left[ \begin{array}{ccc}
    0.253  &       0  & -0.072\\
         0  & 0.246   &  0\\
   -0.072  &       0  &  0.096
\end{array} \right] \ \mathrm{kg\ m}^{2}$\\

\multicolumn{3}{l}{\textbf{F}\emph{ront wheel F}}\\
radius & \emph{$r_\mathrm{F}$} & 0.344 m\\
mass & \emph{$m_\mathrm{F}$} & 2.02 kg \\
mass moments of inertia & \emph{$(I_{\mathrm{F}xx},\
I_{\mathrm{F}yy})$} &
(0.090,\ 0.149) $\mathrm{kg\ m}^2$ \\
\hline
\end{tabular}
}
\label{tab:par}
\end{table*}

\bibliographystyle{acm}
\bibliography{bicycle}
\end{document}