### 2010-10-19 papers/BMD2010/PhysicalParametersPoster/poster.tex
poster draft added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LaTeX poster template
% Created by Nathaniel Johnston
% August 2009
% http://www.nathanieljohnston.com/index.php/2009/08/latex-poster-template/
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[final]{beamer}
\usepackage[scale=1.24]{beamerposter}
\usepackage{graphicx}			% allows us to import images

%-----------------------------------------------------------
% Define the column width and poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% The separation I chose is 0.024 and I want 4 columns
% Then set onecolwid to be (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be 2*onecolwid + sepwid = 0.464
%-----------------------------------------------------------

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{1189mm}
\setlength{\paperheight}{841mm}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\onecolwid}{0.1712\paperwidth}
\setlength{\twocolwid}{0.3664\paperwidth}
\setlength{\threecolwid}{0.5616\paperwidth}
\setlength{\topmargin}{-0.5in}
\usetheme{confposter}

%-----------------------------------------------------------
% Define colours (see beamerthemeconfposter.sty to change these colour definitions)
%-----------------------------------------------------------

\setbeamercolor{block title}{fg=ngreen,bg=white}
\setbeamercolor{block body}{fg=black,bg=white}
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70}
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10}

%-----------------------------------------------------------
% Name and authors of poster/paper/research
%-----------------------------------------------------------

\title{Accurate Measurement of Bicycle Parameters}
\author{Jason K. Moore$^{*}$, Mont Hubbard$^{*}$, A. L. Schwab$^\dag$,
            J. D. G. Kooijman$^\dag$}
\institute
{
\centering
\begin{tabular}{cc}
$^*$ Mechanical and Aerospace Engineering, University of California, Davis
\quad
& $^\dag$ Laboratory for Engineering Mechanics, Delft University of Technology\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu
& e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com
\end{tabular}
}
%-----------------------------------------------------------
% Start the poster itself
%-----------------------------------------------------------
% The \rmfamily command is used frequently throughout the poster to force a serif font to be used for the body text
% Serif font is better for small text, sans-serif font is better for headers (for readability reasons)
%-----------------------------------------------------------

\begin{document}
\frame{
%\begin{frame}[t]

% Divide the frame into columns
\begin{columns}[t] % the [t] option aligns the column's content at the top

% Empty side column for spacing
\begin{column}{\sepwid}\end{column}	% empty spacer column

% First column with content
\begin{column}{\onecolwid}
  \begin{block}{Introduction}
    \rmfamily{
    Accurate measurements of a bicycle's physical parameters are required for
    realistic dynamic simulations and analysis. For the most basic models the
    geometry, mass, mass location and mass distributions of each rigid body
    must be measured. In particular, we are concerned with the measurement of
    the minimal bicycle
    parameters required for the benchmark bicycle presented
    in~\cite{Meijaard2007}. A set of 25 parameters is used to populate the
    model.
    The experimental methods described herein are based primarily on
    the work done in \cite{Kooijman2006} and \cite{Roland1971} but have been
    refined for improved accuracy and methodology. We measured the
    characteristics of six different bicycles, two of which were set up in two
    different configurations. We provide a total of eight different parameter sets
    that can be used with, but not limited to, the benchmark bicycle model. The
    accuracy of all the measurements are presented along with a comparison of
    the linear characteristics of the eight bicycles.
    }
  \end{block}
  \vskip2ex

  % Info about the model
  \begin{alertblock}{Whipple Bicycle Model}
    \rmfamily{
    The linear unforced two degree-of-freedom, $\mathbf{q}$
    = [steer and roll], linear Whipple model takes the form:
    \small{
    \begin{equation}
        \mathbf{M\ddot{q}}
        +v\mathbf{C}_1\mathbf{\dot{q}}
        +\left[g\mathbf{K}_0
        +v^2\mathbf{K}_2\right]\mathbf{q}
        =0
        \label{eq:canonical}
    \end{equation}
    }
    Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
    matrices are combinations of 25 of the bicycle's physical parameters that
    include the geometry, mass, mass location and mass distribution of the four
    rigid bodies.
    }
  \end{alertblock}

\end{column}

% Spacer column
\begin{column}{\sepwid}\end{column} % empty spacer column

% Second Column
\begin{column}{\twocolwid}
  %\setbeamercolor{block title}{fg=red,bg=white}%frame color
  %\setbeamercolor{block body}{fg=black,bg=white}%body color

  \begin{block}{Accuracy}
    \rmfamily{
    Following the vain of \cite{Roland1971} we used error propagation theory to
    calculate accuracy of the 25 benchmark bicycle parameters. We estimate
    the standard deviation of the raw measurements. If $x$ is a parameter and is a function of
    the raw measurements, $u,v,\ldots$, then $x$ is a random variable defined as
    $x=f(u,v,\ldots)$. The sample variance of $x$ is defined as
    \begin{equation}
        s_x^2 = s_u^2\left(\frac{\partial x}{\partial u}\right)^2 +
                s_v^2\left(\frac{\partial x}{\partial v}\right)^2 +
                2s_{uv}\left(\frac{\partial x}{\partial u}\right)\left(\frac{\partial x}{\partial v}\right)
                + \ldots
        \label{eqn:variance}
    \end{equation}
    where $s_*$ is the variance and $s_{**}$ is the covariance. If $u$ and $v$ are uncorrelated then $s_{uv}=0$ but the cross correlations
    must be taken into account otherwise.
    }
  \end{block}
  \begin{columns}[t, totalwidth=\twocolwid]
      \begin{column}{\onecolwid}
          \begin{block}{Bicycles}
            \rmfamily{
            The six bicycles, chosen for both variety and convenience, are as follows:
            \emph{Batavus Browser}, a Dutch style city bicycle measured with and without
            instrumentation; \emph{Batavus Stratos
            Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
            Dutch style city bicycle with a suspended fork; \emph{Gary Fisher Mountain
            Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
            track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
            road bicycle measured in two configurations, the second with the fork rotated
            in the headtube 180 degrees for larger trail.
            \\
            \begin{center}
                \begin{tabular}{ccc}
                    \includegraphics[width=0.333\onecolwid]{../../../images/browserIns_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/crescendo_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/fisher_sub.jpg}
                    \\
                    \includegraphics[width=0.333\onecolwid]{../../../images/pista_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/stratos_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/yellow_sub.jpg}
                \end{tabular}
            \end{center}
            }
          \end{block}
    \end{column}

    \begin{column}{\onecolwid}

        \begin{block}{Measurements}
            \rmfamily{
            We attempted to directly measure the trail, wheelbase and wheel
            diameters with various techniques. We measured the mass centers of
            the four bodies (fork, frame, and wheels) by hanging the bodies by
            their mass centers. The moments of inertia were estimated by
            hanging the bodies as both torsional and compound pendulums and
            measuring the period of oscillation which is correlated to the
            inertia.
            }
            \begin{center}
              \includegraphics[width=0.90\onecolwid]{../../../figures/angles.pdf}
            \end{center}
        \end{block}

    \end{column}
\end{columns}

\end{column}

% Spacer column
\begin{column}{\sepwid}\end{column} % empty spacer column

% The fourth column
\begin{column}{\onecolwid}
  \begin{alertblock}{Bike Eigenanalysis}
    \rmfamily{
    Distinct variation is apparent among the bicycles' eigenvalues. Frequency,
    damping, and the stable speed range can be compared to the physical
    differences of the bicycles.
    }
    \\
    \begin{center}
      \includegraphics[width=.9\onecolwid]{../../../plots/Bike/eig_plot.pdf}
    \end{center}
    \rmfamily{
    Less variation is seen among the bicycles when the mass and inertia of a
    rigid rider is added.
    }
    \\
    \begin{center}
      \includegraphics[width=.9\onecolwid]{../../../plots/BikeRider/eig_plot.pdf}
    \end{center}
  \end{alertblock}

\end{column}

% Spacer column
\begin{column}{\sepwid}\end{column} % empty spacer column

\begin{column}{\onecolwid}
  % Frequency Response
  \begin{block}{Bike Frequency Response}
    \rmfamily{
    The steer torque to roll angle Bode plot reveals up to 10 dB variation
    among the bikes along the frequencies shown.
    \\
    \begin{center}
      \includegraphics[width=.9\onecolwid]{../../../plots/Bike/Bode/Tdel2phi.pdf}
    \end{center}
    }
  \end{block}

  % Frequency Response
  \begin{block}{Bike/Rider Frequency Response}
    \centering
    \includegraphics[width=.9\onecolwid]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
  \end{block}

  % References
  \begin{block}{References}
    \small{\rmfamily{
    \bibliographystyle{plain}
    \bibliography{bicycle}
    }}
  \end{block}
\end{column}
%\vskip2.5ex

% Last spacer column
\begin{column}{\sepwid}\end{column}			% empty spacer column

\end{columns}
}
%\end{frame}
\end{document}